---
title: "Machine Learning Hw6"
author: "Ekta Chaudhary"
date: "12/05/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r}
library(ISLR)
library(factoextra) 
library(gridExtra)
library(corrplot)
library(RColorBrewer) 
library(gplots) 
library(igraph)
library(jpeg)
library(imager)
```

# Data 
```{r}
data("USArrests")
seed = 1
usarrests = scale(USArrests)
```

# Cluster Analysis

We perform hierarchical clustering on the states using the USArrests data in the ISLR
package. For each of the 50 states in the United States, the dataset contains the number
of arrests per 100,000 residents for each of three crimes: Assault, Murder, and Rape. The
dataset also contains the percent of the population in each state living in urban areas,
UrbanPop. The four variables will be used as features for clustering.

## Question a) Using hierarchical clustering with complete linkage and Euclidean distance, cluster the states.

```{r}
set.seed(seed)
hc_noscale = hclust(dist(USArrests, method = 'euclidean'), method = 'complete')
```

## Question b) Cut the dendrogram at a height that results in three distinct clusters. Which states belong to which clusters?

```{r}
hc_3_noscale = cutree(hc_noscale, 3) 
```

## States in the first cluster

```{r}
USArrests[hc_3_noscale == 1,]
```

### States in the 2nd Cluster
```{r}
USArrests[hc_3_noscale == 2,]
```

### States in the 3rd Cluster

```{r}
USArrests[hc_3_noscale == 3,]
```

Visualization of the dendrogram and showing graphically which states belong to which clusters

```{r}
fviz_dend(hc_noscale, k = 3,
          cex = 0.5,  
          palette = "lancet", 
          color_labels_by_k = TRUE, 
          rect = TRUE, 
          rect_fill = TRUE, 
          rect_border = "jco", 
          type = 'rectangle',
          labels_track_height = 60.5
          )
```

## Question c) Hierarchically cluster the states using complete linkage and Euclidean distance, after scaling the variables to have standard deviation one. 

```{r}
set.seed(seed)
hc_scaled = hclust(dist(usarrests, method = 'euclidean'), method = 'complete')
```

```{r}
fviz_dend(hc_scaled, k = 3,
          cex = 0.6,  
          palette = "lancet", 
          color_labels_by_k = TRUE, 
          rect = TRUE, 
          rect_fill = TRUE, 
          rect_border = "jco", 
          type = 'rectangle',
          labels_track_height = 1.5
          )
```


## Question d) What effect does scaling the variables have on the hierarchical clustering obtained? In your opinion, should the variables be scaled before the inter-observation dissimilarities are computed?

```{r}
# number of states in each cluster when scaling
table(cutree(hc_scaled, 3))
# number of states in each cluster when not scaling
table(cutree(hc_noscale, 3))
```

# PCA

```{r}
img <- readJPEG('cat.jpg')
plot(load.image('cat.jpg'))
dim(img)
r <- img[,,1]
g <- img[,,2]
b <- img[,,3]
img.r.pca <- prcomp(r, center = FALSE)
img.g.pca <- prcomp(g, center = FALSE)
img.b.pca <- prcomp(b, center = FALSE)
rgb.pca <- list(img.r.pca, img.g.pca, img.b.pca)
# Approximate X with XV_kV_k^T
compress <- function(pr, k)
{
  compressed.img <- pr$x[,1:k] %*% t(pr$rotation[,1:k])
  compressed.img
}
# Using first 20 PCs
pca20 <- sapply(rgb.pca, compress, k = 20, simplify = "array")
writeJPEG(pca20, "pca20.jpg")
# Try to increase the number of PCs!
plot(load.image('pca20.jpg'))
```
```{r}
plot(load.image('Panda.jpg'))
img <- readJPEG('Panda.jpg')
dim(img)
r <- img[,,1]
g <- img[,,2]
b <- img[,,3]
img.r.pca <- prcomp(r, center = FALSE)
img.g.pca <- prcomp(g, center = FALSE)
img.b.pca <- prcomp(b, center = FALSE)
rgb.pca <- list(img.r.pca, img.g.pca, img.b.pca)
# Approximate X with XV_kV_k^T
compress <- function(pr, k)
{
  compressed.img <- pr$x[,1:k] %*% t(pr$rotation[,1:k])
  compressed.img
}
# Using first PC
p1 <- sapply(rgb.pca, compress, k = 1, simplify = "array")
writeJPEG(p1, "parrot1.jpg")
plot(load.image('parrot1.jpg'))
# Using first 10 PCs
p10 <- sapply(rgb.pca, compress, k = 10, simplify = "array")
writeJPEG(p10, "parrot10.jpg")
plot(load.image('parrot10.jpg'))
# Using first 50 PCs
p50 <- sapply(rgb.pca, compress, k = 50, simplify = "array")
writeJPEG(p50, "parrot50.jpg")
plot(load.image('parrot50.jpg')) 
# Using first 100 PCs
p100 <- sapply(rgb.pca, compress, k = 100, simplify = "array")
writeJPEG(p100, "parrot100.jpg")
plot(load.image('parrot100.jpg')) 
# Using first 200 PCs
p200 <- sapply(rgb.pca, compress, k = 200, simplify = "array")
writeJPEG(p200, "parrot200.jpg")
plot(load.image('parrot200.jpg')) 
```